Rice yield improvements through plant breeding are offset by inherent yield declines over time

### 2018-02-23 output/breed_paper_revised.Rnw
Commenting on Kents 2014 pub.

>>>>

\citet{mckenzie2014} explored the yield of rice in this same system
during a similar time period using similar data as these here, but
concluded that genetic rice yield potential has increased over this
time period at a rate of 25 kg\tsups{-1} ha\tsups{-1}
year\tsups{-1}. There are several explanations for how this could
be. First, \citet{mckenzie2014} did not account for yield decreases
over time. As shown here, when these decreases are not taken into about, the
estimated increase in yield over time is estimated to be approximately
the same as that found by \citet{mckenzie2014}. Second,
\citet{mckenzie2014} concluded that there was no trend in the yield
of the major cultivars over time, yet this interpretation of the
results of a statistical test is flawed. Failing to reject the
hypothesis that of a trend in yield does not equate to there being no
trend. Third, it is not surprising that no trend in yield over time
was detected. \citet{mckenzie2014} used data from several locations, yet analyzed
those data using a simple t-test. By using this simple analysis,
\citet{mckenzie2014} ignore systematic errors related to differences between
observations, including different management systems, growing
environments, and crop timing between locations. By contrast, this
investigation used a single, well managed system and analyzed the data
using a mixed effects model which fully accounts for non-independence
between similar observations. In doing this, the analysis presented
here makes fewer assumptions and more fully captures the structure of
the data. Hence, the estimates presented here are likely more
reliable.
